---
title: "2^k Factorial Designs"
subtitle: "STA6246"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Catalyst of Chemical Process

-   An experimenter wanted to study the Yield of a chemical process

-   The engineer decides to test 2 reactant concentrations (15 vs 25) and 2 catalysts (1 vs 2).

-   3 replicates were run

### Data description

```{r}
library(tidyverse)
library(gtsummary)
# 2^2 Factorial design
# 2^2 Data Yield in a chemical process

Y=c(28,36,18,31,25,32,19,30,27,32,23,29)
A=gl(n=2,k=1,length = 4,labels = c("L","H"))
B=gl(n=2,k=2,length = 4,labels = c("Lo","Hi"))

# block
bl=gl(n=3,k=4,length = 12,labels=c(1,2,3))
# data
mdata=data.frame(Y,A,B)

# ANOVA
# EA= 8.33
# EB = -5
model1 <- aov(Y ~ A+B, data = mdata)
summary(model1)

```

### Computing Effects 

Estimate the effects from the regression

#### The contrasts 

```{r}
getOption('contrasts')

```

#### Orthogonal - effect coding - half-effects

```{r}
options(contrasts=c("contr.sum","contr.poly"))
contrasts(A)
m2=lm(Y ~ A*B, data = mdata)
coef(m2)

# change reference level
co=coef(lm(Y ~ relevel(A, ref = "H")*relevel(B, ref = "Hi"),data = mdata))

# Effects
Eff=2*co[2:4]
names(Eff) = c("A","B","AB")
Eff
```

#### Reference coding -  effects

It works only with no interactions

```{r}

options (contrasts = c("contr.treatment", "contr.poly"))
contrasts(A)

m3 = lm(Y ~ A+ B ,data=mdata)
#By default The intercept: overall mean - (EffectA+ EffectB)/2
# and the regression coefficients are the Effect Estimates

Eff2=coef(m3)[2:3]
names(Eff2) <- c("A","B") 
Eff2

# Graphics
interaction.plot(x.factor = mdata$A, 
                 trace.factor = mdata$B, 
                 response = mdata$Y, 
                 fun = mean, 
                 type = "b", 
                 col=c("black" ,"red"),
                 fixed = T,
                 pch = c(19,17))



```

## Etch Rate data

An etch process on a single wafer plasma etching tool. The design factors:

-   The gap between the electrodes (A)

-   The gas flow (B)

-   RF power applied to the cathode. (C)

-   Each factor is run at two levels

-   The design is replicated twice n=2.

```{r}
Etch=c(550,669,633,642,1037,749,1075,729,604,650,601,635,1052,868,1063,860)
A=gl(n=2,k=1,length = 16,labels = c("L","H"))
B=gl(n=2,k=2,length = 16,labels = c("L","H"))
C=gl(n=2,k=4,length = 16,labels = c("L","H"))
df=tibble(Etch,A,B,C)


model1 <- aov(Etch ~ A*B*C, data = df)
summary(model1)


# Get estimates
getOption("contrasts")
options(contrasts=c("contr.sum","contr.poly")) # provide half-effects
co=coef(lm(Etch ~ relevel(A, ref = "H")*relevel(B, ref = "H")*relevel(C, ref = "H")))
# Effects
Eff=2*co[2:8]
names(Eff) <- c("A","B","C","AB","AC","BC","ABC") 
Eff

qq=qqnorm(Eff,cex=1,ylim = c(-200,300),xlim=c(-3,1.5))
qqline(Eff)
text(qq$x,qq$y,names(Eff),pos=1,cex=1,col="red")



```

-   The factors A, C, and interactions AC are the most important terms.
