---
title: "STA6246 - DOE"
subtitle: "ANOVA Examples"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Etching Process Experiment

The goal of this experiment is to compare 4 Radio Frequency (RF) power to obtain the desired etching rate.

### Data description

```{r}
library(tidyverse)
library(gtsummary)
# enter ER
Etch_Rate=c(575,542,530,539,570,565,593,590,579,610,600,651,610,637,629,725,700,715,685,710)
# enter RF
RF=gl(n = 4, k = 5, labels = c("160","180","200","220"))
# put together data in a tibble
etch_exp=tibble(Etch_Rate,RF)

etch_exp %>%
  ggplot(aes(x=RF,y=Etch_Rate)) +
  geom_point()+
  theme_bw()+
  labs(x="Radio Frequency Power (W)",
       y="Etch Rate (unit)")


etch_exp %>%
  tbl_summary(by=RF)
```

-   The etching rate seems to increase with the RF power.

-   The variability seems to be similar among the groups (powers)

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor
results = aov(Etch_Rate ~ RF,data = etch_exp)
summary(results)
tab_model(results)

# Individual CIs - need to add the CI for the reference level
confint(results)
```

The ANOVA was rejected (p-value $< 0.001$). Therefore, the data is compatible with the means difference in the population. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)

```

The residuals plots do not show any major departures from the ANOVA assumptions.

##### Multiple Comparison

Considering 5% level of significance.

```{r}

# Multiple Comparison
pairwise.t.test(Etch_Rate,RF,p.adjust.method = "bonferroni")

#Tukey Test 
THSD=TukeyHSD(results)
plot(THSD,las=1)

# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"RF",console=TRUE,group = F)
```

**Conclusion**:

The Radio Frequency Power setting does affect the etching rate in the etching process. The higher the RF power the higher the etch rate.

## Peak Discharge Data

An Engineer is interested in determining whether **4 different methods** of estimating flood flow frequency produce equivalent estimates of **peak discharge** when applied to the same watershed. **Six** replications were run. Therefore. we have 6\*4= **24 runs.**

```{r}
pd=c(0.34,0.12,1.23,0.7,1.75,0.12,0.91,2.94,2.14,2.36,2.86,4.55,6.31,8.37,9.75,6.09,9.82,7.24,17.15,11.82,10.95,17.20,14.35,16.82)

methods=gl(n = 4, k = 6, length = 24, labels = c("a","b","c","d"))

peak_discharge_exp = tibble(pd,methods)
```

### Data description

```{r}

peak_discharge_exp %>%
  ggplot(aes(x=methods,y=pd)) +
  geom_point()+
  theme_bw()+
  labs(x="Methods",
       y="Peak Discharge")


peak_discharge_exp %>%
  tbl_summary(by=methods,
              statistic = list(pd ~ "{mean} ({sd}); [{min}, {max}]") )

```

-   The Method "a" yields to the higher peak discharge values compared to other methods

-   The variability seems to be a bit differnet among the groups (methods)

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor
results = aov(pd ~ methods,data = peak_discharge_exp)
summary(results)
tab_model(results)

# Individual CIs - need to add the CI for the reference level
confint(results)
```

The ANOVA was rejected (p-value $< 0.001$). Therefore, the data is compatible with the means difference in the population. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)

```

The residuals plots do not show any major departures from the ANOVA assumptions, except the variance constant assumption.

##### Cox-Box Transformation for stabilizing variance

```{r}
library(car)
PT=powerTransform(results)

peak_discharge_exp = peak_discharge_exp %>%
  mutate(pdT= 2*(sqrt(pd)-1)) 


# ANOVA one factor
resultsT = aov(pdT ~ methods,data = peak_discharge_exp)
summary(resultsT)
tab_model(resultsT,results)

# residuals plots
par(mfrow=c(2,2))
plot(resultsT)
```

The p-value \<0.001 so the ANOVA null hypothesis is rejected. Data supports evidence of peak discharge mean difference among the methods.

##### Multiple Comparison

Considering 5% level of significance.

```{r}

# Multiple Comparison
pairwise.t.test(pd,methods,p.adjust.method = "bonferroni")

#Tukey Test 
THSD=TukeyHSD(results)
plot(THSD,las=1)

# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"methods",console=TRUE,group = F)
```

**Conclusion**:

The Methods do affect the peak discharge values. The method "d" provides the highest peak discharge. Also, methods "a" and "b" seems to result in similar mean peak difference (p=0.266; 95%CI -0.93 to 4.76 using Tukey's test).

## Fabric Strength and Looms - Random ANOVA

A textile company weaves a fabric on a large number of looms.

The engineer suspects that in addition to the usual variation in **strength** within the samples of the same loom, there may be significant variations in strength between looms.

The engineer selected **a=4** looms at random and **n=4** replicates.

#### Data description

```{r}
strength=c(98,97,99,96,91,90,93,92,96,95,97,95,95,96,99,98)
looms=gl(n = 4, k = 4, labels = c("lm1","lm2","lm3","lm4"))

strength_loom = tibble(strength,looms)

strength_loom %>%
  ggplot(aes(x=looms,y=strength)) +
  geom_boxplot()+
  theme_bw()+
  labs(x="Looms",
       y="Strength")
```

#### ANOVA - To test the variance component hypothesis

```{r}
# to calculate the mean squares
aRE=aov(strength ~ looms, data = strength_loom)
summary(aRE)
confint(aRE)
par(mfrow=c(2,2))
plot(aRE)
```

The results from ANOVA indicates significant difference in the variability due to in the looms. We want to estimate variance components:

##### Moments Method

```{r}
# Estimate the variance components - method of moments
summary(aRE)
s2= 1.9
s2tau= (29.729 - 1.896)/4
s2y = s2 + s2tau
ICC = s2tau/s2y

```

##### MLE method

```{r}
library(lme4)
# Fit the Random Effect One-way ANOVA
anovaRE <- lmer(strength ~ 1 + (1 | looms))
summary(anovaRE)
# Confidence Intervals Variance Components and Overall Mean
confint(anovaRE,oldNames=FALSE)
# Estimates the random effect tau_i 
ranef(anovaRE)
# Check the residuals
par(mfrow = c(1, 2))
plot(anovaRE)

#ICC
library(irr)
t(matrix(strength, nrow =  4))
icc(t(matrix(strength, nrow =  4)))
```

> The **conclusion**: The strength variability is significantly different from loom to loom (p \<0.001). The variance component are:
>
> -   The variance due to the looms is 6.958
>
> -   The variance of the error: 1.896
>
> -   The ICC is 0.786 (95% CI \[0.39 to 0.98\])
>
> We can conclude that the 78.6% of variability in the strength is due to the difference among the looms.

## Cardiovascular and Chocolate Study

The goal is to investigate the effect of consuming chocolate on cardiovascular health (Plasma Antioxidants from Chocolate, Nature, 2003). 3 types of chocolates:

-   100 g of dark chocolate : DC

-   100 g of dark chocolate with 200mL of full fat milk: DCMK

-   200 g of milk chocolate: MK

12 subjects. The total antioxidant capacity of their blood plasma was measured after 1 hour of consumption of one of the above level.

```{r}
DC    =c(118.8 ,	122.6 ,	115.6 ,	113.6 ,	119.5 ,	115.9 ,	115.8 ,	115.1 ,	116.9 ,	115.4 ,	115.6 ,	107.9)

DCMK  =c(105.4 ,	101.1 ,	102.7 , 97.1, 	101.9 ,98.9 ,	100.0,99.8,102.6,100.9,104.5,93.5)

MC    =c(102.1 ,	105.8 ,	99.6  ,	102.7 ,98.8 ,	100.9 ,	102.8 ,98.7 ,94.7 ,97.8 ,99.7 ,98.6)

subjects = gl(n=12,k=1,length = 36,labels = (1:12))

cardio_chocolate = tibble(DC,DCMK,MC) %>%
  pivot_longer(cols = everything(),
               names_to = "group",
               values_to = "antioxidantCap") %>%
  mutate(subjects = gl(n=12,k=3,length = 36,labels = (1:12))
)

```

```{r}
cardio_chocolate %>%
  ggplot(aes(x=group,y=antioxidantCap))+
  geom_boxplot()+
  theme_bw()


res = aov(antioxidantCap~group+subjects,data=cardio_chocolate)
summary(res)

# variance components
library(lme4)
anovaRE <- lmer(antioxidantCap ~ group + (1 | subjects), data = cardio_chocolate)
summary(anovaRE)

# Confidence Intervals Variance Components and Overall Mean
confint(anovaRE,oldNames=FALSE)

```
