---
title: "Balanced Incomplete Block Design"
subtitle: "STA6246"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Introduction

-   In certain experiments using randomized block designs, we may not be able to run all the treatment combinations in each block.

-   Situations like this usually occur because of shortages of experimental apparatus or facilities or the physical size of the block.

-   For example, in the **vascular graft experiment**, suppose that each batch of material is only large enough to accommodate testing three extrusion pressures. Therefore, each pressure cannot be tested in each batch.

-   **When all treatment comparisons are equally important,** the treatment combinations used in each block should be selected in a balanced manner, so that any pair of treatments occur together the same number of times as any other pair. This is a balanced incomplete block design (BIBD).

## Reaction Time with 4 Catalysts

|        | Catalyst1 | Catalyst2 | Catalyst3 | Catalyst4 |
|--------|-----------|-----------|-----------|-----------|
| Batch1 | 73        | NA        | 73        | 75        |
| Batch2 | 74        | 75        | 75        | NA        |
| Batch3 | NA        | 67        | 68        | 72        |
| Batch4 | 74        | 72        | NA        | 75        |

-   An experimenter wanted to study the chemical **reaction time** of with 4 **catalysts**.

-   Each formulation is mixed from a **batch** of raw material that only enough for 3 reaction.

-   We cannot "afford" to do a complete design.

How can we generate a BIBD, let's use the function below:

```{r}
library(multcomp)
# 4 is the number of treatments (catalysts)
# 3 is the number of experimental unit per block
# 4 choose 3 will be the number of blocks
combn(x=4,m=3)
```

Every column is a block. This is called **unreduced** **BIBD.** Now, in practice we can offer to do all blocks.

```{r}
library(ibd)
des.bibd <- bibd(v = 6, b = 10, r = 5, k = 3, lambda = 2) 
## arguments of function bibd are:
## - v: number of treatments
## - b: number of blocks
## - r: number of replicates (across all blocks)
## - k: number of experimental units per block
## - lambda: lambda

des.bibd$design ## here, blocks are given by *rows*
```

### Data

```{r}
library(tidyverse)
library(gtsummary)

# BIBD EXAMPLE: 
TimeR=c(73,74,NA,71,NA,75,67,72,73,75,68,NA,75,NA,72,75)
Catalyst= gl(n = 4,k=4,labels = c(1:4))
Batch= gl(n = 4,k=1,length = 16,c(1:4))

mydata = tibble(TimeR, Catalyst, Batch)

TimeR_result = aov(TimeR ~ Batch+ Catalyst, data = mydata)

library(ibd)

des.bibd = bibd(v = 4, b =  4, r = 3, k = 3, lambda = 2) 
## v: number of treatments
## b: number of blocks
## k: number of units per block
## r: number of replications
##
des.bibd$design

## Test if the Design is BIBD
library(crossdes)
tab <- xtabs(TimeR~ Catalyst+Batch, data = mydata)
d <- t(apply(tab,1, function(x) (1:4)[x != 0]))
isGYD(d)



aov.ibd(TimeR ~ Batch+Catalyst, data = mydata)


fit <- aov(TimeR ~ Batch + Catalyst , data = mydata)
drop1(fit, test = "F")









rocket=tibble(BurnRate,Bt,Op,Trts,Tests)

rocket %>%
  ggplot(aes(x=Trts,y=BurnRate,group=Trts)) +
  geom_point()+
  stat_summary(aes(group=1),fun = mean, geom = "line", col="red",linewidth=2)+
  stat_summary(aes(label=..y..,group=1),fun = mean, geom = "text", col="blue",size=10)+
  theme_bw()+
  labs(x="Formualtions",
       y="Burn Rate")

rocket %>% 
  select(BurnRate,Trts) %>%
  tbl_summary(by=Trts,
              statistic = list(all_continuous() ~ "{mean} ({sd})"))
```

-   The mean burning rate seems to vary acrross the formulations.

-   The variability seems to vary a bit across the formulations

-   The Formulation D provides the highest mean burning rate and B the lowest.

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor and one blocking factor
results = aov(BurnRate ~ Trts + Bt + Op + Tests ,data = rocket)
summary(results)
tab_model(results)
```

```{r}

# Individual CIs 
confint(results)
```

The ANOVA was rejected (p-value $= 0.007$). Therefore, the data is compatible with the means difference in the burn rate across the 5 formulations. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)
```

We could consider a transformation because the Scale-Location plots is showing a trend.

##### Transformation

```{r}
library(car)
l=powerTransform(results)
l$roundlam
```

Since $\lambda$ was estimated to be $0$ no need to a transformation.

##### Multiple Comparison

Considering 5% level of significance.

```{r}

#Tukey Test 
THSD=TukeyHSD(results,which = "Trts")
plot(THSD,las=1)
THSD
# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"Trts",console=TRUE,group = F)
```

**Conclusion**:

-   Tukey's Test indicates:

    -   Significant differences between:

        -   B and D;

        -   A and B; and

        -   C and D.

-   Fisher LSD Test indicates:

    -   Significant differences between:

        -   B and **D**;

        -   A and B;

        -   C and **D**;

        -   B and E;

        -   A and C.

Formulation D led to the highest mean burning rate and it is statistical significant different from formations B and C.

Formulation D, A, and E resulted in no significant differences.

#### 
