---
title: "Graeco-Latin Square Design"
subtitle: "STA6246"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Introduction

-   The Graeco-Latin Square Design (GLSD) allows us to eliminate **three** nuisance factors (blocking).

-   Each factor has p levels

-   We can control the effect of these blocking factors by grouping the experimental units into blocks.

-   This is an extension of the Latin Square Design.

## Rocket Propellant Experiment

-   An experimenter wanted to study the effects of 5 different formulations of a rocket propellant on the observed **burning rate**.

-   Each formulation is mixed from a **batch** of raw material.

-   Each formulation is prepared by one **operator** for each batch.

-   Each formulation can be assembly tested using 5 different **tests**.

-   $p=5$ levels for each factor

$p^2= 25$ runs or observations.

### Data description

```{r}
library(tidyverse)
library(gtsummary)
# enter ER
BurnRate=c(24,17,18,26,22,20,24,38,31,30,19,30,26,26,20,24,27,27,23,29,24,36,21,22,31)
Bt=gl(n = 5, k = 1,length = 25, labels = c(1:5))
Op=gl(n = 5, k = 5,length = 25, labels = c(1:5))
Trts=c("A","B","C","D","E",
       "B","C","D","E","A",
       "C","D","E","A","B",
       "D","E","A","B","C",
       "E","A","B","C","D")

# ADDITIONAL FACTOR - THAT IS THE TEST ASSEMBLIES
Tests=c("Aa","Ba","Ca","Da","Ea",
        "Da","Ea","Aa","Ba","Ca",
        "Ba","Ca","Da","Ea","Aa",
        "Ea","Aa","Ba","Ca","Da",
        "Ca","Da","Ea","Aa","Ba")

rocket=tibble(BurnRate,Bt,Op,Trts,Tests)

head(rocket)

rocket %>%
  ggplot(aes(x=Trts,y=BurnRate,group=Trts)) +
  geom_point()+
  stat_summary(aes(group=1),
               fun = mean, 
               geom = "line", 
               col="red",
               linewidth=2)+
  stat_summary(aes(label=..y..,group=1),
               fun = mean, 
               geom = "text", 
               col="blue",
               size=10)+
  theme_bw()+
  labs(x="Formualtions",
       y="Burn Rate")

rocket %>% 
  select(BurnRate,Trts) %>%
  tbl_summary(by=Trts,
              statistic = list(all_continuous() ~ "{mean} ({sd})"))
```

-   The mean burning rate seems to vary acrross the formulations.

-   The variability seems to vary a bit across the formulations

-   The Formulation D provides the highest mean burning rate and B the lowest.

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor and one blocking factor
results = aov(BurnRate ~ Trts + Bt + Op + Tests ,data = rocket)
summary(results)
tab_model(results)
```

```{r}

# Individual CIs 
confint(results)
```

The ANOVA was rejected (p-value $= 0.007$). Therefore, the data is compatible with the means difference in the burn rate across the 5 formulations. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)
```

We could consider a transformation because the Scale-Location plots is showing a trend.

##### Transformation

Since $\lambda$ was estimated to be $0$ we need a $\log$ transformation.

```{r}
library(car)
l=powerTransform(results)
l$roundlam
## do the transofrmation
Tburn =log(BurnRate)
## Refit ANOVA
rano=aov(Tburn ~ Bt + Op + Trts + Tests, data=rocket)
## Summary
summary(rano)

par(mfrow=c(2,2))
plot(rano)
```

##### Multiple Comparison

Considering 5% level of significance.

```{r}
#Tukey Test 
THSD=TukeyHSD(results,which = "Trts")
plot(THSD,las=1)
THSD
# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"Trts",console=TRUE,group = F)
```

**Conclusion**:

-   Tukey's Test indicates:

    -   Significant differences between:

        -   B and D;

        -   A and B; and

        -   C and D.

-   Fisher LSD Test indicates:

    -   Significant differences between:

        -   B and **D**;

        -   A and B;

        -   C and **D**;

        -   B and E;

        -   A and C.

Formulation D led to the highest mean burning rate and it is statistical significant different from formations B and C.

Formulation D, A, and E resulted in no significant differences.

#### 
