---
title: "STA6246 - DOE"
subtitle: "LSD - ANOVA Examples"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Introduction

-   The Latin Square Design (LSD) allows us to eliminate two nuisance factors (blocking).
-   Each factor has p levels
-   We can control the effect of these blocking factors by grouping the experimental units into blocks.
-   We consider a square wit p rows and p columns.

We can geenrate latin square in R as follows:

```{r}
library(magic)
# generate r by r latin square design
rlatin(5)
```

## Rocket Propellant Experiment

-   An experimenter wanted to study the effects of 5 different formulations of a rocket propellant on the observed burning rate.

-   Each formulation is mixed from a batch of raw material.

-   Each formulation is prepared by one operator for each batch.

-   $p=5$ levels for each factor

$p^2= 25$ runs or observations.

### Data description

```{r}
library(tidyverse)
library(gtsummary)
# enter ER
BurnRate=c(24,17,18,26,22,20,24,38,31,30,19,30,26,26,20,24,27,27,23,29,24,36,21,22,31)
Bt=gl(n = 5, k = 1,length = 25, labels = c(1:5))
Op=gl(n = 5, k = 5,length = 25, labels = c(1:5))
Trts=c("A","B","C","D","E",
       "B","C","D","E","A",
       "C","D","E","A","B",
       "D","E","A","B","C",
       "E","A","B","C","D")

rocket=tibble(BurnRate,Bt,Op,Trts)

rocket %>%
  ggplot(aes(x=Trts,y=BurnRate)) +
  geom_point()+
  theme_bw()+
  labs(x="Formualtions",
       y="Burn Rate")


rocket %>%
  tbl_summary(by=Trts)
```

-   The formulations seem to impact the burnig rate

-   The variability seems to be similar across the formulations

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor and one blocking factor
results = aov(BurnRate ~ Trts + Bt + Op - 1 ,data = rocket)
summary(results)
tab_model(results)
```

```{r}

# Individual CIs - need to add the CI for the reference level
confint(results)
```

The ANOVA was rejected (p-value $= 0.003$). Therefore, the data is compatible with the means difference in the burn rate across the 5 fomulations. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)
```

We could consider a transformation because the Scale-Location plots is showing a trend.

##### Transformation

```{r}
library(car)
l=powerTransform(results)
l$roundlam
Tburn =log(BurnRate)
rano=aov(Tburn ~ Bt + Op + Trts)
summary(rano)

boxplot(Tburn~Trts)
par(mfrow=c(2,2))
plot(rano)
```

##### Multiple Comparison

Considering 5% level of significance.

```{r}

#Tukey Test 
THSD=TukeyHSD(results,which = "Trts")
plot(THSD,las=1)

# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"Trts",console=TRUE,group = F)
```

**Conclusion**:

Formation D led to the higher mean burning rate and it is statistical significant different from formation C (p=0.025; 95% CI 0.82 to 14) and B (p=0.004; 95% CI 3.016 to 16.18).

#### 
