---
title: "STA6246 - DOE"
subtitle: "RCBD - ANOVA Examples"
author: "Dr. Cohen "
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
toc: true
---

## Vascular Graft Experiment

A medical manufacturer wants to investigate **a=4** different levels of **extrusion pressure** on flicks (defects) by using randomized complete block design.

The experimenter considers the batches of resin (raw material) as **blocks** **b=6**.

**N= 24** runs or observations.

### Data description

```{r}
library(tidyverse)
library(gtsummary)
# enter ER
dataV=read.csv(file = "../datasets/VascularData.csv",row.names = 1)
dataV$Press=factor(dataV$Press, labels=c("8500","8700","8900","9100"))
dataV$Batch=as.factor(dataV$Batch)


dataV %>%
  ggplot(aes(x=Press,y=Yield)) +
  geom_point()+
  theme_bw()+
  labs(x="Extrusion pressure (PSI)",
       y="Yiled (%)")


dataV %>%
  tbl_summary(by=Press)
```

-   The yield seems to decrease with the extrusion pressure.

-   The variability seems to be similar among the groups (extrusion pressure)

### ANOVA in R

##### Fitting

```{r}
library(sjPlot)

# ANOVA one factor and one blocking factor
results = aov(Yield ~ Press + Batch ,data = dataV)
summary(results)
```

```{tab_model(results)}

# Individual CIs - need to add the CI for the reference level
confint(results)
```

The ANOVA was rejected (p-value $= 0.0019$). Therefore, the data is compatible with the means difference in the yields population among the 4 extrusion pressures. Next, we will diagnose the model.

##### Model Adequacy

```{r}
par(mfrow=c(2,2))
plot(results)

```

We could consider a transformation because the Scale-Location plots is showing a trend.

##### Multiple Comparison

Considering 5% level of significance.

```{r}

#Tukey Test 
THSD=TukeyHSD(results,which = "Press")
plot(THSD,las=1)

# Fisher LSD Test
library(agricolae)
L=LSD.test(results,"Press",console=TRUE,group = F)
```

**Conclusion**:

The Extrusion Pressure setting does affect the Yield (defects) in the Vascular Graft Experiment. The Lower the Pressure the higher the Yield.

The 9100 pressure was significantly different from 8500 and 8700 pressures. However, the 9100 pressure statistically resulted in no difference in the yield compared to 8900 pressure.

#### Random Effects RCBD

```{r}
library(lme4)
anovaRE <- lmer(Yield ~ 1 + Press + (1 | Batch), data = dataV)
summary(anovaRE)
# Confidence Intervals Variance Components and Overall Mean
confint(anovaRE,oldNames=FALSE)
# Estimates the random effect
ranef(anovaRE)
anova(anovaRE)
icc = 7.781/(7.781+7.326)
icc
```

There is a significant effect of the variability coming from the batch to batch.
